---
title: 'Sentiment Analysis of Game of Thrones by G.R.R. Martin'
author: "Marius Henningsen, Malte Volquartzen And Soroosh Sabri"
date: "`r format(Sys.time(), '%d %B %Y, %H:%M')`"
output: html_document
---

First we load the packages we will use for the analysis including packages for text mining and visualisation

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse)
library(here)

# For text mining:

library(pdftools)
library(tidytext)
library(textdata) 
library(ggwordcloud)
```


### We get the Game of Thrones pdf from the data folder:
```{r get-document}
GOT_path <- here("data","GOT.pdf")
GOT_path

# We make sure the text is recognized as an pdf text in R and look up a single random page to see if it loaded correctly.
GOT_text <- pdf_text(GOT_path)
# We choose to look up page 9
GOT_p9 <- GOT_text[9]
GOT_p9

```

The "pdf_text()" command returns a vector of strings, one for each page of the pdf. We turn it into a dataframe so we can mess with it in tidyverse style and keep track of the pages. We use stringr::str_split() to break the pages up into individual lines. Each line of the pdf is concluded with a backslash-n, so we split on this. We will also add a line number in addition to the page number."

### Some wrangling:

- First We will split up pages into separate lines (separated by `\n`) using `stringr::str_split()`
- Second We will Unnest into regular columns using `tidyr::unnest()`
- Then We will remove leading/trailing white space with `stringr::str_trim()`

```{r split-lines}
# We split pages into separate lines

GOT_df <- data.frame(GOT_text) %>% 
  mutate(text_full = str_split(GOT_text, pattern = '\n')) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full)) 

GOT_df

```

Now each line, on each page, is its own row, with extra starting & trailing spaces removed. 

### Get the tokens (which means individual words) in tidy format

I will use `tidytext::unnest_tokens()` (which pulls from the `tokenizer`) package, to split columns into tokens. We are interested in *words*, so that's the token we'll use:

```{r tokenize}
GOT_tokens <- GOT_df %>% 
  unnest_tokens(word, text_full)

GOT_tokens
# See how this differs from `GOT_df`
# Each word has its own row!
```

I then arrange the words from the pdf in a tibble arrange them after how frequently they occur
```{r count-words}
GOT_wc <- GOT_tokens %>% 
  count(word) %>% 
  arrange(-n)
GOT_wc
```

A lot of stop words show up that need to be removed.

### Remove stop words:

I will *remove* stop words using `tidyr::anti_join()`:
```{r stopwords}
GOT_stop <- GOT_tokens %>% 
  anti_join(stop_words) %>% 
  select(-GOT_text)

GOT_stop
```

Now We check the counts again: 
```{r count-words2}
GOT_swc <- GOT_stop %>% 
  count(word) %>% 
  arrange(-n)
GOT_swc

```

We can see that titles like Lord and ser and names of the main characters like Jon and Ned are some of the words occurring most frequently

I then get rid of all the numbers (non-text) in `GOT_stop`?
```{r skip-numbers}

GOT_no_numeric <- GOT_stop %>% 
  filter(is.na(as.numeric(word)))
GOT_no_numeric


```

### We will now make a word cloud of the words in GOT 

```{r wordcloud-prep}

#I check how many unique words occur in GOT excluding stop words
length(unique(GOT_no_numeric$word))

# There are almost 11209 unique words excluding stop words and numbers

# We don't want to include all 11209 words in a word cloud, so We filter to only include the top 100 most frequent
GOT_top100 <- GOT_no_numeric %>% 
  count(word) %>% 
  arrange(-n) %>% 
  head(100)
GOT_top100
```

I then make a word cloud of the top 100 most frequent words.

```{r wordcloud}
GOT_cloud <- ggplot(data = GOT_top100, aes(label = word)) +
  geom_text_wordcloud() +
  theme_minimal()

GOT_cloud
```

I then customize it a bit to make the word cloud more effective:
```{r wordcloud-pro}
ggplot(data = GOT_top100, aes(label = word, size = n)) +
  geom_text_wordcloud_area(aes(color = n), shape = "triangle") +
  scale_size_area(max_size = 12) +
  scale_color_gradientn(colors = c("darkgreen","blue","red")) +
  theme_minimal()
```
I have now made an easily readable word cloud showing the most frequent words like Lord, Ser, Jon, Ned and so on.
Now to the sentiment analysis. To get rid of titles and names we could make a list of our own stopwords. But for now We will continue with the sentiment analysis.

### Sentiment analysis

Game of Thrones is grim-dark fantasy which often has a tone, style, or setting that is particularly bleak, amoral, and violent. 
I therefore expect that negative emotions will dominate the text rather than positive emotions.

In my sentiment analysis We will look at three general-purpose lexicons which are

  -  AFINN from Finn Årup Nielsen,
  -  bing from Bing Liu and collaborators, and
  -  nrc from Saif Mohammad and Peter Turney


"afinn": Words ranked from -5 (very negative) to +5 (very positive)
```{r afinn}
get_sentiments(lexicon = "afinn")

# We look at the some of the positive words to see if it works:
afinn_pos <- get_sentiments("afinn") %>% 
  filter(value %in% c(3,4,5))
afinn_pos
```

bing: binary, "positive" or "negative"
```{r bing}
get_sentiments(lexicon = "bing")

```



Now We will look at nrc:
```{r nrc}
get_sentiments(lexicon = "nrc")

```

I will now do sentiment analysis on the GOT text data using afinn and nrc. 


### Sentiment analysis with afinn: 

First We bind words in `GOT_stop` to `afinn` lexicon:
```{r bind-afinn}
GOT_afinn <- GOT_stop %>% 
  inner_join(get_sentiments("afinn"))
GOT_afinn
```

I then make a visualisation of how frequently each sentiment ranking occur in the pdf. 
This will give us an impression of the tone of the language.
```{r count-afinn}
GOT_afinn_hist <- GOT_afinn %>% 
  count(value)

# Plot them: 
ggplot(data = GOT_afinn_hist, aes(x = value, y = n)) +
  geom_col(aes(fill = value)) +
  theme_bw()
```

So we can see that there are more negative sentiment rankings in the pdf, a lot of them being of the value -2 

I then investigate some of the words in a bit more depth:
```{r afinn-2}
# We look at the words with the value 2
GOT_afinn2 <- GOT_afinn %>% 
  filter(value == 2)

GOT_afinn2
```

```{r afinn-2-more}
# We then check the unique 2-score words:
unique(GOT_afinn2$word)

# We count & plot them
GOT_afinn2_n <- GOT_afinn2 %>% 
  count(word, sort = TRUE) %>% 
  mutate(word = fct_reorder(factor(word), n))


ggplot(data = GOT_afinn2_n, aes(x = word, y = n)) +
  geom_col() +
  coord_flip() +
  theme(text = element_text(size = 12))+
  theme_bw()

ggsave(here("figures","GOT_afinn2_n.png"), 
       height = 17, 
       width = 5)

# We can see the word honor topping the list of 2-score words. First of all tells us that honor is a significant theme in the book, and without the proper context it can be a bit hard to sat whether the book has a positive or negative approach to honor
```

I summarize the Afinn sentiment analysis and find out what the mean and median score of the sentiment score 
```{r summarize-afinn}
GOT_summary <- GOT_afinn %>% 
  summarize(
    mean_score = mean(value),
    median_score = median(value)
  )
GOT_summary 
```

The mean and median indicate slightly negative overall sentiments based on the AFINN lexicon. Game of Thrones is considered a grim-dark fantasy book, so it makes sense for the sentiments to be slightly negative.

### We then use the NRC lexicon for sentiment analysis

We can use the NRC lexicon to start "binning" text by the feelings they're typically associated with. As above, we'll use inner_join() to combine the GOT non-stopword text with the nrc lexicon: 

```{r bind-bing}
GOT_nrc <- GOT_stop %>% 
  inner_join(get_sentiments("nrc"))

GOT_nrc
```

That will exclude some of the words in the text. We check which are excluded using `anti_join()`:

```{r check-exclusions}
GOT_exclude <- GOT_stop %>% 
  anti_join(get_sentiments("nrc"))

# View(GOT_exclude)

GOT_exclude

# Count to find the most excluded:
GOT_exclude_n <- GOT_exclude %>% 
  count(word, sort = TRUE)

GOT_exclude_n

head(GOT_exclude_n)
```

One of the words which are excluded by NRC is 'dead', a word with a very negative sentiment. It's noteworthy that NRC doesn't include this, because it makes NRC weaker as a tool for sentiment analysis.

Now We find counts of NRC sentiments in GOT and plot them: 
```{r count-bing}
GOT_nrc_n <- GOT_nrc %>% 
  count(sentiment, sort = TRUE)

# And plot them:

ggplot(data = GOT_nrc_n, aes(x = sentiment, y = n)) +
  geom_col(aes(fill = sentiment))+
  theme_bw()
```

I then make a visualisation of the counts by sentiment **and** word.
```{r count-nrc}
GOT_nrc_n5 <- GOT_nrc %>% 
  count(word,sentiment, sort = TRUE) %>% 
  group_by(sentiment) %>% 
  top_n(5) %>% 
  ungroup()

GOT_nrc_gg <- ggplot(data = GOT_nrc_n5, aes(x = reorder(word,n), y = n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, ncol = 2, scales = "free") +
  coord_flip() +
  theme_minimal() +
  labs(x = "Word", y = "count")

# Show it
GOT_nrc_gg

# Save it
ggsave(plot = GOT_nrc_gg, 
       here("figures","GOT_nrc_sentiment.png"), 
       height = 8, 
       width = 5)

```
Here we can see weird examples showing up like 'stone' and 'words' being associated with anger and 'mother' being associated with anticipation. A name like Bran is also associated with disgust because NRC sees it as a normal word and not a name. All these things would have to be kept in mind and edited to make sense for the given text.
